---
title: "Homework 2"
output: html_document
---

```{r include = FALSE, echo = FALSE}
# Libraries needed
library(tidyverse)
library(patchwork)
library(palmerpenguins)

# global datasets and variables needed
mpg <- ggplot2::mpg
penguins <- palmerpenguins::penguins
```

# Task 1

Using the mpg dataset, compare and contrast the following plots. These plots should be displayed next to each other.

This two plots compare the `city miles per gallon` (`cty`) with the `highway miles per gallon` (`hwy`) with two different displaying methods:
`geom_point` and `geom_count`.
```{r}
# create both plots
plot1 <- ggplot(data = mpg, aes(x = cty, y = hwy)) + geom_point()
plot2 <- ggplot(data = mpg, aes(x = cty, y = hwy)) + geom_count()

# combine them with the `patchwork` library
plot1 + plot2
```

# Task 2
Using the `palmerpenguins::penguins` dataset, explain whether there is any problem with the graph below, and eventually propose a solution.

```{r}
# inital plot
ggplot(data = penguins, aes(fill = island, x = species)) + geom_bar(aes(y = after_stat(prop)))
```
This plot seems only to show the islands inhabited by the different penguin species.
We are not able to see how big each population is and do not get any further information.

Proposed solution: Make a plot to show the distribution of one species on the islands with a barchart
```{r}
ggplot(penguins, aes(species, fill = island)) +
  geom_bar() +
  geom_text(aes(label = after_stat(count), y = after_stat(count)), stat = "count", vjust = 1.5)
```

# Task 3
To visualyze the distribution of the weight of diamonds with respect to their color, we have the following options:

## Facet the histogram on weight
```{r}
ggplot(diamonds, aes(x = carat)) + geom_histogram(bins = 30) + facet_wrap(~ color, ncol = 2)
```

## Superpose the density plots
```{r}
ggplot(diamonds, aes(x = carat, fill = color)) + geom_density()
```

## Display boxplots
```{r}
ggplot(diamonds, aes(x = color, y = carat)) + geom_boxplot()
```

## Use violin plots
```{r}
ggplot(diamonds, aes(x = color, y = carat)) + geom_violin()
```