---
title: "Homework 2"
output:
  html_document:
    df_print: paged
---

```{r include = FALSE, echo = FALSE}
# Libraries needed
library(nycflights13)
library(ggplot2)
library(ggridges)
library(readxl)
library(dlookr)
library(tidyverse)

```

# Task 1

Using a single pipeline and `nycflights13::flights`, figure out the number of days on which the temperature
was at least 30°C at JFK-Airport, broken down by month. Do these months correspond to what you expected?

```{r}
flights |>
  merge(weather, by = intersect(names(flights), names(weather))) |>
  filter(origin=="JFK") |>
  filter(temp >= 86.00) |>  # 86°F == 30°C, temp is in F
  distinct(year, month, day) |>
  group_by(month) |>
  count(month)
```

The result shows us 2 days in July, 14 in June, 2 in August and 1 in September.
This sem to match the average temperature per month in NYC.

# Task 2
Using a single pipeline and `nycflights13::flights`,
figure out December flights missing `arr_time` and get all other variables the flights missing `arr_time` are also missing.
What might these rows represent?

```{r}
flights |>
  filter((month == 12) & (is.na(arr_time))) |>
  summarize(across(everything(), ~ sum(is.na(.))))
```

The correlated variables are `dep_time`, `dep_delay`, `arr_time`, `arr_delay`, `tailnum`, and `air_time`.
- `dep_time` and `dep_delay` have the same count and could tell us, that the flight did not happen in 2013.
- `arr_time`, `arr_delay`, and `air_time` have the same count and could tell us, that the departed flights did not arrive in 2013
- `tailnum` could tell us, that the flights where planned but the plane could not be assigned

For a more detailed and possible better answer we would have to investigate the concrete dates. Only then we could validate the three proposed cases.
Overall it looks like the `NA` variables are due to the year change and could be resolved with the data for 2014.

# Task 3
Select all variables from `nycflights13::weather` except those between year and hour.
Use this selection to solve the tasks `a)`, `b)` and `c)` below.
For the diagrams in `a)` and `b)`, label the months with their full name.
For the diagram in `c)`, use number to label the months.

```{r}
sel_weather <- weather |>
  select(!(year:hour))
```
The selection seems to be wrong and not suited for the following tasks.
In the following tasks we need to use the excluded variables.
But for the sake of this homework, lets do this.

## a)
Display the distribution of the temperature faceted on months.

```{r}
# since we need the month variable (which we exclueded before in the selection), we use here the original weather dataset
adj_weather <- weather
adj_weather$month_name <- month.name[adj_weather$month]

ggplot(adj_weather, aes(x = temp)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  facet_wrap(~ month_name, ncol = 3) +
  labs(title = "Temperature Distribution by Month", x = "Temperature (°C)", y = "Frequency")
```

## b)
Use `ggridges::geom_density_ridges()` to get a plot similar to the one in a).

```{r}
# since we need the month variable (which we exclueded before in the selection), we use here the original weather dataset
adj_weather <- weather
adj_weather$month_name <- month.name[adj_weather$month]

ggplot(adj_weather, aes(x = temp, y = month_name)) +
  geom_density_ridges(fill = "blue", color = "black") +
  labs(title = "Temperature Distribution by Month", x = "Temperature (°C)", y = "Month") +
  scale_y_discrete(labels = month.name) +
  theme_ridges() +
  theme(legend.position = "none")
```

## c)
Aggregate your data into average monthly temperature, including their standard deviation.
Use the aggregated data and geom_line() to display monthly average temperature
surrounded by the ±2 · s / √n upper and lower 95% confidence lines,
where s is a monthly standard deviation and n the √ number of observation per month.

```{r}
aggregated_weather <- adj_weather |>
  group_by(month) |>
  filter(!is.na(temp)) |>
  summarize(
    mean_temp = mean(temp),
    std_dev = sd(temp),
    n = n()
  ) |>
  mutate(
    upper_limit = mean_temp + 2 * (std_dev / sqrt(n)),
    lower_limit = mean_temp - 2 * (std_dev / sqrt(n))
  )

ggplot(aggregated_weather, aes(x = month, y = mean_temp)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower_limit, ymax = upper_limit), fill = "blue", alpha = 0.3) +
  labs(title = "Monthly Average Temperature with 95% Confidence Intervals",
       x = "Month",
       y = "Average Temperature (°C)")
```

# Task 4
[Data_HW3.xlsx](./Data_HW3.xlsx) has in the sheet `small sample`
some data collected on BFH-W 3rd semester students a couple of years ago.
Most of the variables are self explaining.

- `foot` gives foot size measured in cm without shoes.
- `hair` is the length of respondent’s longest hair measured in cm.
- `transport` is the main method of transport used by the respondent to get to the BFH.
- `cash` (CHF) gives how much money cash in CHF the student had with her/him at the time she/he was filling the survey.

- Import this data into R and inspect it. Describe any problem you notice and propose a solution for it.
- Save the cleaned data as `.csv` in your R-project directory.

```{r results = FALSE}
data_hw3 <- read_excel("./Data_HW3.xlsx")

# as I have learnd in SBD2: lets use diagnose from dlookr
overview(data_hw3)
diagnose(data_hw3)

# diagnose shows us that we have some NA values inside the dataset. since they are just a few, we just remove them.
clean_hw3 <- data_hw3 |>
        drop_na()

# lets write the new dataset to output
write.csv(clean_hw3, file = "./cleaned_data_Homework_3.csv")
```