---
title: "Homework 4 - Exploratory Data Analysis"
subtitle: "Part 1 - Group work"
author: "Severin Clauss"
date: Nov. 19, 2023
output: pdf_document
fontsize: 12pt
---

```{r include = FALSE, echo = FALSE}
# Install libraries needed
libraries <- c('readxl', 'ggplot2', 'forcats', 'tidyverse', 'patchwork')

lapply(libraries, function(x) if (!(x %in% installed.packages())) {
  install.packages(x)
})

lapply(libraries, library, quietly = TRUE, character.only = TRUE)

if (basename(getwd()) != "homework-4") {
  setwd("./homework-4")
}


# read the dataset
dataImprot <- read_excel('./Data_HW4.xlsx')
data <- dataImprot
```

# Exploratory Data Analysis
*`Data_HW4.xlsx` contains data collected from BFH-W 3rd semester students. See `Data_Collection.pdf`. Use your EDA-skills to explore and summarize your findings.*

For this task, we decided to take first a dive into the data of the transportation.

The first question in our mind was: Which transportation is used the most?
And next, we would like to know the mean distance per transportation?
```{r}
# what is the most used transportaition?
viz_transport <- ggplot(data = data, aes(x = fct_infreq(transport), fill = transport)) +
        geom_bar() +
        coord_polar() +
        labs(title = 'Frequency of transportation used', y = '', x = '')

# now we can see, that the most come by train followed by bus and then car.
# so what are the distances here for the ones by car, bus or train?
viz_distance <- data |>
        aggregate(distance ~ transport, mean) |>
        ggplot(aes(
                x = reorder(transport, distance, decreasing = TRUE),
                y = distance,
                fill = transport
        )) +
        geom_bar(stat = "identity", color = "white", show.legend = FALSE) +
        labs(title = 'Mean distance by transportation', x = '')

# plot both visualizations at once to have a nice comparison
viz_transport + viz_distance
```

Now we can see that the ones who travel by train have also the longest distance to travel.
This seems to make sense since you most likely use the train to travel bigger distances.
This also tells us that the majority of the students live outside Bern
and rely therefore on transportation like train and car.

\pagebreak
Next we ask our self if transportation has any influence on sleep?

```{r fig.width=20, fig.height=10}
# lets check first for outliers
viz_box_sleep <- ggplot(data, aes(x = sleep)) +
  geom_boxplot() +
  labs(title = 'Distribution of sleep')

# Since the outliers seem realistic, we keep them in our analysis.
viz_freq_sleep <- data |>
  aggregate(sleep ~ transport, mean) |>
  ggplot(aes(
    x = reorder(transport, sleep, decreasing = TRUE),
    y = sleep,
    fill = transport
  )) +
  geom_bar(stat = "identity", color = "white", show.legend = FALSE) +
  labs(title = 'Mean sleep over transportation',
       x = "Transport",
       y = "Sleep")

viz_box_sleep + viz_freq_sleep
```

The ones who walk have most likely the shortest way and can therefore sleep the longest.
Interesting is the public transportation vs bicycle.
It seems that the ones with public transportation can sleep longer than the ones by bicycle.
This does not match with the distance traveled.

\pagebreak
So let us check the observation with boxplots to check what is happening here:
```{r fig.height=3}
ggplot(data, aes(x = transport, y = sleep, fill = transport)) +
  geom_boxplot() +
  coord_flip() +
  labs(title = "Sleep distribution over transportation",
       x = "Sleep",
       y = "Transport")
```

The sleep distribution on bicycle seems strongly left skewed.
Therefore, we conclude that this is not representable and needs to be taken with a big grain of salt.

\pagebreak
After checking out the transport, distance and sleep, we decided to check out the gender-specific parts in this dataset:

First, we want to know the distribution of gender and class in the dataset. We figured out, that there are six classes represented:
```{r fig.height=3}
# There are 6 classes ranging from 29 to 114 students
viz_class1 <- data %>%
  ggplot(aes(x=class)) +
  geom_bar() +
  labs(title = 'Count per class', x = 'class', y = 'count')

#Gender
data %>%
  group_by(gender) %>%
  count()

#Plot class and gender
viz_class2 <- data %>%
  ggplot(aes(x=class, fill = gender)) +
  geom_bar() +
  labs(title = 'Count per class split by gender', x = 'class', y = 'count')

viz_class1 + viz_class2
```

There are significantly more male people in the dataset and just one NB person

Let us understand the height difference between the gender:
```{r}
# we take only the gender Female and Male cause non binary is just one Person
data_gender <- data |>
  filter(gender %in% c("Female", "Male"))

# eliminate the outlieres
Data_HW4_filtered <- data_gender %>%
  group_by(gender) %>%
  mutate(Q1 = quantile(height, 0.25),
         Q3 = quantile(height, 0.75),
         IQR = Q3 - Q1,
         lower_limit = Q1 - 1.5 * IQR,
         upper_limit = Q3 + 1.5 * IQR) %>%
  filter(height >= lower_limit & height <= upper_limit)

# Boxplot height gender
ggplot(Data_HW4_filtered, aes(x = gender, y = height, fill = gender)) +
  geom_boxplot() +
  coord_flip() +
  labs(title = 'Distribution hight per gender', x = 'height', y = '')
```

We now understand that at least in the representation of this dataset,
the female observations are generally smaller than the male counterparts.

\pagebreak
Let us also have a look at the age distribution regarding gender:
```{r fig.height=3}
# Fix dob and make it a date
data_dob <- data %>%
  mutate(dob = as.Date(dob))

#People born before 1900 are most likely not alive anymore
data_dob %>%
  filter(dob > "1970-01-01") %>%
  ggplot(aes(x=dob, fill = gender)) +
  geom_histogram() + labs(title = 'Distribution of age and gender', x = 'Date of birth', y = '')
```
