---
title: "Homework 4"
author: "Severin Clauss"
date: Nov. 19, 2023
output: pdf_document
fontsize: 10pt
---

```{r include = FALSE, echo = FALSE}
# Install libraries needed
libraries <- c('readxl', 'ggplot2', 'forcats', 'tidyverse', 'patchwork')

lapply(libraries, function(x) if (!(x %in% installed.packages())) {
  install.packages(x)
})

lapply(libraries, library, quietly = TRUE, character.only = TRUE)

if (basename(getwd()) != "homework-4") {
  setwd("./homework-4")
}


# read the dataset
dataImprot <- read_excel('./Data_HW4.xlsx')
data <- dataImprot
```

# Task 1
*`Data_HW4.xlsx` contains data collected from BFH-W 3rd semester students. See `Data_Collection.pdf`. Use your EDA-skills to explore and summarize your findings.*

For this task, we decided to take first a dive into the data of the transportation. The first question in our mind was: Which transportation is used the most? And next, we would like to know the mean distance per transportation?
```{r}
# what is the most used transportaition?
viz_transport <- ggplot(data = data, aes(x = fct_infreq(transport), fill = transport)) +
        geom_bar() +
        coord_polar() +
        labs(title = 'Frequency of transportation used', y = '', x = '')

# now we can see, that the most come by train followed by bus and then car.
# so what are the distances here for the ones by car, bus or train?
viz_distance <- data |>
        aggregate(distance ~ transport, mean) |>
        ggplot(aes(
                x = reorder(transport, distance, decreasing = TRUE),
                y = distance,
                fill = transport
        )) +
        geom_bar(stat = "identity", color = "white", show.legend = FALSE) +
        labs(title = 'Mean distance by transportation', x = '')

# plot both visualizations at once to have a nice comparison
viz_transport + viz_distance
```

# Task 2