---
title: "Homework 4 - Exploratory Data Analysis"
subtitle: "Part 2 - Individually"
author: "Severin Clauss"
date: Nov. 19, 2023
output: pdf_document
fontsize: 12pt
---

```{r include = FALSE, echo = FALSE}
# Install libraries needed
libraries <- c('tidyverse', 'ggplot2')

lapply(libraries, function(x) if (!(x %in% installed.packages())) {
  install.packages(x)
})

lapply(libraries, library, quietly = TRUE, character.only = TRUE)

if (basename(getwd()) != "homework-4") {
  setwd("./homework-4")
}
```

*`anscombe` illustrates how graphs are essential to good statistical analysis. These data are from four samples of eleven observations each. The index of x and y represents the corresponding sample. For example, the first observation in sample 1 has the x-value 10 and the y-value 8.04, and the first observation in sample 2 has the x-value 10 and the y-value 9.14.*

# A)
*Explain briefly why this dataset is not tidy. In a single pipeline, transform it to get a tidy-dataset with exactly three variables: sample, x and y. Do not output the tidy-dataset, but save it for future use.*

The dataset is not tidy because it violates the principles of tidy data, which include:

- Each variable forms a column.
- Each observation forms a row.
- Each type of observational unit forms a table.

```{r}
anscombe_data_tidy <- anscombe |>
  pivot_longer(cols = everything(),
               names_to = c(".value", "sample"),
               names_pattern = "([xy])(\\d)") |>
  arrange(sample)
```

\pagebreak
# B)
*In a single pipeline, use the tidy-data from [a)](#a) to get a scatterplot that includes a line of best fit for y on x. The sample should be distinguished by color and shape. Use only the shapes 0, 1, 2 and 5.*

```{r}
anscombe_data_tidy |>
        ggplot(aes(x = x, y = y, color = factor(sample), shape = factor(sample))) +
          geom_point() +
          geom_smooth(method = "lm", se = FALSE) +
          scale_shape_manual(values = c(0, 1, 2, 5)) +
          labs(title = "Scatterplot with Line of Best Fit",
             x = "X Values",
             y = "Y Values",
             color = "Sample",
             shape = "Sample")
```

\pagebreak
# C)
*In a single pipeline, summarize the tidy-data from [a)](#a) to get the means, standard deviations and correlation coeﬃcients.*

```{r}
anscombe_data_tidy |>
        group_by(sample) |>
        summarise(
                mean_x = mean(x),
                mean_y = mean(y),
                sd_x = sd(x),
                sd_y = sd(y),
                corr_xy = cor(x, y)
        )
```

# D)
*In a single pipeline, use facet_grid() to display the scatter plots per sample (including lines of the best fit), and summarize the main message on your plot.*

```{r}
anscombe_data_tidy |>
  ggplot(aes(x = x, y = y, color = factor(sample))) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  facet_grid(. ~ sample) +
  labs(title = "Scatterplots with Lines of Best Fit per Sample",
       x = "X Values",
       y = "Y Values",
       color = "Sample") +
  theme_minimal()
```

The main message of this plot is to emphasize how different datasets can have similar summary statistics (mean, standard deviation, correlation) but exhibit different patterns when visualized.